---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# realarea

<!-- badges: start -->
<!-- badges: end -->

The goal of realarea is to provide helpers to determine the actual area covered and the calculated area vs the "real" geographic area represented in a projected raster. 


## Installation

You can install the development version of realarea like so:

``` r
remotes::install_github("hypertidy/realarea")
```

## Example

We have a shapefile and we want to assess suitability of a given map projection. 

```{r example}
library(realarea)

luxshp <- system.file("ex/lux.shp", package="terra", mustWork = TRUE)
lux <- terra::vect(luxshp)
```

With `crs_grid()` we can create a suitable raster grid from that vector data set. 

By default we just get a nice grid from it in the native crs. 

```{r crs_grid}
crs_grid(lux)
```

But specify an actual map projection and we can get some more benefit, and with 'res' we can get exactly what we want: a raster from which we can calculate real-world area (with `terra::expanse()`) and compare that with the nominal cell size which is `prod(res(<x>))`. 


```{r crs}
crs <- "EPSG:23032"
crs_grid(lux, crs = crs)

luxutm1000 <- crs_grid(lux, crs = crs, res = 1000)
luxaea1000 <- crs_grid(lux, crs = "+proj=aea +lon_0=6 +lat_0=50 +lat_1=50.1 +lat_2=49.4", res = 1000)
```

What is the total area of the input polygon?

What do we get from the two projections?

```{r area}
library(terra)

polyarea <- sum(expanse(lux))
sqrt(polyarea)
utmarea <- sum(values(mask(cellSize(luxutm1000), luxutm1000)), na.rm = TRUE)
sqrt(utmarea)


## this one is closest to the polygon area
aeaarea <- sum(values(mask(cellSize(luxaea1000), luxaea1000)), na.rm = TRUE)
sqrt(aeaarea)

plot(cellSize(luxutm1000)/prod(res(luxutm1000)))

plot(cellSize(luxaea1000)/prod(res(luxaea1000)))

```


So, unsurprsingly the local Albers Equal Area Conic projection is a better choice than UTM, but
do we really care about that level of discrepancy?  Probably not, but you can't have a single rule 
that always works, it depends where, how long how wide, on the projection, what you need to measure, 
and what you are actually doing. :)

When folks look for a crs, they often restrict themselves only to codes (e.g. EPSG) that are pre-defined, that can work ok for nations and particular well-mapped regions, but there's no hard rule, you can define your own projection for particular purposes. 


What about other fun projections?


```{r vicgrid}
library(terra)
library(sf)
p <- vect(silicate::inlandwaters[5:6, ])

utm <- crs_grid(p, "EPSG:32755")

plot(cellSize(utm)/prod(res(utm)))
plot(project(p, "EPSG:32755"), add = TRUE)

vicgrid <- crs_grid(p, "EPSG:7899")

plot(cellSize(vicgrid)/prod(res(vicgrid)))
plot(project(p, "EPSG:7899"), add = TRUE)

```

What if we had a region like the Albers national projection used by GA? 

```{r wateven,eval=FALSE, include=FALSE}
share <- strsplit(system("whereis proj", intern  = TRUE), " ")[[1]][4]
con <- RSQLite::dbConnect(RSQLite::SQLite(), file.path(share, "proj.db"))
library(dbplyr)
library(dplyr)
DBI::dbListTables(con)
library(stringr)
albex <- tbl(con, "extent") |> filter(code == 3577) |> transmute(xmin = west_lon, xmax = east_lon, ymin = south_lat, ymax = north_lat) |> collect() |> unlist()
albex
```

So, what does that look like? 

```{r vaster}
albex <- c(112.85,153.69,-43.7,-9.86)
bdy <- reproj::reproj_xy(vaster::vaster_boundary(c(32, 32), albex), 
                         "EPSG:3577", source = "EPSG:4326")

plot(bdy, asp = 1)

```

```{r albersgrid}
library(terra)
p <- terra::vect(matrix(albex[c(1, 1, 2, 2, 1, 
             3, 4, 4, 3, 3)], ncol = 2), type = "polygons", crs = "EPSG:4326")
p <- terra::densify(p, 1, flat = TRUE)  ## beware densify, we don't want great circles ...

grd <- crs_grid(p, "EPSG:3577")
terra::plot(grd); plot(project(p, "EPSG:3577"), add = TRUE)


plot(cellSize(grd)/prod(res(grd)))
oz <- vect(sds::CGAZ(), query = sds::CGAZ_sql("Australia"))
plot(project(oz, "EPSG:3577"), add = TRUE)
```

Let's expand that out to Australia's broader remit, this takes a bit of special handling and is teaching me something I needed to figure out ... WIP

```{r ozex}
ozex <- c(55, 164, -90, -6)
p <- terra::vect(matrix(ozex[c(1, 1, 2, 2, 1, 
             3, 4, 4, 3, 3)], ncol = 2), type = "polygons", crs = "EPSG:4326")
p <- terra::densify(p, 1000)
grd <- crs_grid(p, "EPSG:3577")
grd <- crop(grd, c(xmin(grd), xmax(grd), -7.5e6, ymax(grd)))
agrd <- cellSize(grd)
agrd[!agrd > 1e5] <- NA
plot(agrd/prod(res(agrd)))
m <- reproj::reproj_xy(do.call(cbind, maps::map(plot = F)[1:2]), "EPSG:3577", source = "EPSG:4326")
lines(m)
```

## Code of Conduct
  
Please note that the realarea project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
